Validating risk scores

### 2020-12-03 manuscript/manuscript.Rmd
Tweaks VK, formatting fixes for the supplements

We collected data on comorbidities and information about disease progression on daily resolution including breathing support required, oxygen flow, experimental anti-Covid-19 and antimicrobial drugs taken and several laboratory markers (PCR positivity for SARS-CoV-2, C-reactive protein, D-dimer, Interleukin 6, Ferritin, lymphocyte count). Full protocol for data collection is attached in the supplementary material. Due to very low number of patients using extra-corporeal membrane oxygenation (N = `r n_ecmo`) or non-invasive positive pressure ventilation (N = `r n_nippv`) in our sample, we merged those categories with mechanical ventilation.

Not all patients developed pneumonia or other symptoms of Covid-19. All patients received the standard of care including supplemental oxygen and ventilation and antibiotics for bacterial superinfections as determined by the attending physician. Some patients were not indicated for all treatment modalities (especially mechanical ventilation) based on decision of the attending physician and underlying patient condition. We note that the participating sites were not homogeneous in either patient population or treatment protocols. The choice of pharmacological treatment was based on the decision of the attending clinician and its availability.

>>>>

We collected data on comorbidities and information about disease progression on daily resolution including breathing support required, oxygen flow rate, experimental anti-Covid-19 and antimicrobial drugs taken and several laboratory markers (PCR positivity for SARS-CoV-2, C-reactive protein, D-dimer, Interleukin 6, Ferritin, lymphocyte count). Full protocol for data collection is attached in the supplementary material. Due to very low number of patients using extra-corporeal membrane oxygenation (N = `r n_ecmo`) or non-invasive positive pressure ventilation (N = `r n_nippv`) in our sample, we merged those categories with mechanical ventilation.

Not all patients developed pneumonia or other symptoms of Covid-19. All patients received the standard of care which could include supplemental oxygen and ventilation and antibiotics for bacterial superinfections, as determined by the attending physician. Some patients were not indicated for all treatment modalities (especially mechanical ventilation) based on decision of the attending physician and underlying patient condition. We note that the participating sites were not homogeneous in either patient population or treatment protocols. The choice of pharmacological treatment was based on the decision of the attending clinician and its availability.

### 2020-12-03 manuscript/risk_scores.Rmd
Tweaks VK, formatting fixes for the supplements

shi_et_al_compare %>%
  transmute(score = shi, `Shi et al. (admission)` = shi_severe_admission, `Ventilated (admission)` = 1 - our_not_ventilated_admission, `Oxygen (admission)` = 1 - our_not_oxygen_admission, N = n_our_admission)

shi_et_al_compare %>%
  transmute(score = shi, `Shi et al. (hospitalization)` = shi_severe_all, `Ventilated (hospitalization)` = 1 - our_not_ventilated_all, `Oxygen (hospitalization)` = 1 - our_not_oxygen_all, N = n_our_all)

@@

```{r}
age_only_data <- at_admission %>%
  mutate(age_decade = floor(age / 10))
age_auc <-rbind(
  my_auc_summary(alive_30 ~ age, age_only_data),
  my_auc_summary(alive_12 ~ age, age_only_data),
  my_auc_summary(alive_30 ~ age_decade, age_only_data, note = "decade"),
  my_auc_summary(alive_12 ~ age_decade, age_only_data, note = "decade")
)
age_auc

@@

We see that the "age only" baseline performs similarly to the best performing score (Caramelo et al.) all of the other scores - there is some uncertainty prohibiting us to be very sure about which works better. This holds even if we only take the decade of age. Note that the score by (Caramelo et al.) is actually heavily based on the decade of age. Table of the results is shown below as the small differences are not well visible in the plot.

>>>>

shi_et_al_compare %>%
  transmute(score = shi, `Shi et al. (admission)` = shi_severe_admission, `Ventilated (admission)` = 1 - our_not_ventilated_admission, `Oxygen (admission)` = 1 - our_not_oxygen_admission, N = n_our_admission) %>% knitr::kable(booktabs = TRUE)

shi_et_al_compare %>%
  transmute(score = shi, `Shi et al. (hospitalization)` = shi_severe_all, `Ventilated (hospitalization)` = 1 - our_not_ventilated_all, `Oxygen (hospitalization)` = 1 - our_not_oxygen_all, N = n_our_all) %>% knitr::kable(booktabs = TRUE)

@@

```{r}
age_only_data <- at_admission %>%
  mutate(age_decade = floor(age / 10))
age_auc <-rbind(
  my_auc_summary(alive_30 ~ age, age_only_data),
  my_auc_summary(alive_12 ~ age, age_only_data),
  my_auc_summary(alive_30 ~ age_decade, age_only_data, note = "decade"),
  my_auc_summary(alive_12 ~ age_decade, age_only_data, note = "decade")
)
age_auc %>% knitr::kable(booktabs = TRUE)

@@

We see that the "age only" baseline performs similarly to the best performing score (Caramelo et al.) and likely better than all of the other scores - there is however some uncertainty prohibiting us to be very sure about the ordering. This holds even if we only take the decade of age. Note that the score by (Caramelo et al.) is actually heavily based on the decade of age. Table of the results is shown below as the small differences are not well visible in the plot.
